Parallelization of Fast $\ell_1$-Minimization for Face Recognition

### 2011-03-08 paper/main.tex
Working on writing for the implementation and experiments sections

\section{Hardware Parallelism} \label{sec:parallelism} 
In this section we discuss the hardware architectures used in this paper. Aside
from the $\ell_1$-min algorithm, our choice of the architectures plays a
significant role in many practical decisions we make in Section
\ref{sec:alignment} and Section \ref{sec:recognition}.  For CPU implementation,
the system would make use of all the cores in as many CPUs are present on one
motherboard.  For GPU implementation, the system would make the best use of a
single many-core Nvidia GPU chip  on a PCI card (GPUs are commonly shipped with
a single GPU chip per PCI-express card).  Unless otherwise specified, all
implementations utilize single precision floating point datatypes.  Since on-chip
memory (cache) is often orders of magnitude faster than off-chip memory, the sizes
and speeds of on-chip memory (cache) have big performance implications.  We give
a brief overview of the caches that are available in our target architectures,
and defer discussion of their performance implications to sections
\ref{sec:alignment_implementation_cpu} and
\ref{sec:alignment_implementation_gpu}.

@@

\subsection{CPU Implementation}
{\bf AY: A diagram should be used to illustrate how the jobs are distributed on a CPU architecture.}
The CPU ALM algorithm was written in C and compiled with Intel's MKL libraries, which provide
automatic parallelization.

\subsection{GPU Implementation}
{\bf AY: A diagram should be used to illustrate how the jobs are distributed on a GPU architecture.}
The GPU ALM algorithm is written in CUDA and utilizes the CUBLAS libraries.
CUBLAS are the BLAS libraries consisting of all single precision BLAS calls
which are provided by Nvidia and are optimized for GPU performance. In
addition, we coded kernel (GPU) methods for vector-vector operations to replace
operations that would have taken multiple BLAS calls. Because data transfer from the 
RAM to GPU suffers bandwidth limitations, we designed our implementation to transfer the A matrix and measurement vector once.  
The algorithm is similar to the CPU algorithm except that we had to parallelize the code at an
additional level. The CPU portion of the code serves as the conductor, making
calls to GPU functions and directing the flow of the code while the GPU keeps all of the data and runs the computation.

\section{Experiment}
{\bf AY: Describe the benchmark protocol here. Both for simulation and real face data.}

In this section, we present four experiments to benchmark the performance of our algorithm on CPU and GPU platforms. 
To generate a meaningful comparison between different hardware architectures,
in all cases we compare the performance of hardware architectures on a
per-board basis:  For CPU implementations, the benchmark makes use of all of
the cores in as many cpu's are present.  For GPU implementations, the benchmark
makes the best use of the entire GPU chip (most GPU boards have a single GPU
chip).  Unless otherwise specified, all implementations utilize single
precision floating point datatypes.  For all of the experiments, we set the number 
of inner loop iterations at 50 and outer loop iterations at 5000.

@@

\begin{figure}[sim_plot]
\begin{center}
      \includegraphics{../results/random_data/time_vs_matrix_size_constant_tol.svg}
\end{center}
\caption{Speed vs Dimensions of A.  Outer loop tolerance = 0.1.  Inner loop tolerance = $10^{-5}$}
\end{figure}

>>>>

\section{Hardware Parallelism} \label{sec:parallelism} 
In this section we discuss the hardware architectures used in this paper. Aside
from the $\ell_1$-min algorithm, our choice of the architectures plays a
significant role in many practical decisions we make in Section
\ref{sec:alignment} and Section \ref{sec:recognition}.  For CPU implementation,
the system makes use of all the cores in as many CPUs are present on one
motherboard.  For GPU implementation, the system makes the best use of a
single many-core Nvidia GPU chip on a PCI card (GPUs are commonly shipped with
a single GPU chip per PCI-express card).  Unless otherwise specified, all
implementations utilize single precision floating point datatypes.  Since on-chip
memory (cache) is often orders of magnitude faster than off-chip memory, the sizes
and speeds of on-chip memory (cache) have big performance implications.  We give
a brief overview of the caches that are available in our target architectures,
and defer discussion of their performance implications to sections
\ref{sec:alignment_implementation_cpu} and
\ref{sec:alignment_implementation_gpu}.

@@

\subsection{CPU Implementation} Our CPU implementation of Algorithm
\ref{alg:alm_rec} was written in C++.  For operations in the algorithm that map
nicely onto the BLAS API (such as SGEMV, SAXPY {\bf? Victor, fill in here}), we
call Intel's MKL BLAS library, which internally takes advantage of the SSE
units on all available CPU cores. 
{\bf TODO: Our current implementation links against the MKL
library provided internally with matlab; we are in the process of updating the code to link directly against a newer version
of MKL}

For operations that do not map nicely onto the BLAS API, such as the shrinkage
operator, {\bf Argue that these parts are optimized, and/or demonstrate that
performance of these parts is insignificant}.

\subsection{GPU Implementation} Our GPU implementation of Algorithm
\ref{alg:alm_rec} is similar to the CPU implementation described in the
previous section, in that only one problem is solved at a time, and each linear
algebraic operation is parallelized separately.

For most of the operations, calls to MKL CPU BLAS library map directly onto
similar calls in Nvidia's GPU BLAS library (CUBLAS).  The CUBLAS library is
designed to utilize all of the SM's on the GPU.

Operations that do not map well onto the BLAS API (mostly combined
vector-vector operations) were implemented as gpu kernels (in C for CUDA).
Even for operations that could have been implemented via multiple BLAS calls,
this was found to have better performance due to reducede bandwidth and kernel
call overhead {\bf Victor, verify or correct this}.

In order to avoid expensive data transfer across the bus connecting the GPU
card and the CPU motherboard (the PCI express bus), all of the data is
transferred to GPU DRAM once, and all non-trivial tasks are performed on the
GPU on data stored in GPU DRAM.  The sole exception to the above implementation
strategy is that the pseudoinverse of A is still computed on the CPU via the
LAPACK implementation that is part of Intel's MKL. This operation need only be
performed (and results uploaded to GPU DRAM) once before GPU computation
starts. {\bf we should prove justification for A pseudoinverse being cheap,
either here or perhaps later in a pie chart}

\section{Experiments}

In this section, we benchmark the performance of our algorithm on CPU and GPU platforms. 
In order to show how our $\ell_1$ minimization algorithms scale with problem size,
we will begin with benchmarks on synthetic data.  We will then progress to demonstrating 
the speed of our implementations as applied to the face recognition problem.

To generate a meaningful comparison between different hardware architectures,
in all cases we compare the performance of hardware architectures on a
per-board basis:  For CPU implementations, the benchmark makes use of all of
the cores in as many cpu's are present.  For GPU implementations, the benchmark
makes the best use of the entire GPU chip (most GPU boards have a single GPU
chip).  Unless otherwise specified, all implementations utilize single
precision floating point datatypes.

@@

\begin{figure}[sim_plot]
\begin{center}
      \includegraphics{../results/random_data/time_vs_matrix_size_constant_tol}
\end{center}
\caption{Speed vs Dimensions of A.  Outer loop tolerance = 0.1.  Inner loop tolerance = $10^{-5}$}
\end{figure}
